Over-relaxed Sinkhorn--Knopp Algorithm for Regularized Optimal Transport

### 2017-10-20 paper.tex

Note that the local linear rate is also improved if $\theta_0$ is not the optimal choice, see \cite[chapter 5]{ciarlet1982introduction}. By remarking that for any choice of $\theta_0<2$ there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative, one sees that this result applies in particular to our algorithm because eventually, for $\ell$ big enough, one has $\Theta(\gamma^\ell)=\theta_0$.
\begin{corollary}
The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
\end{corollary}
\begin{proof}
For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual} and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0 e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$. Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line. Also, this invariance disappears in the associated primal iterates.

The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative}, even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.

>>>>

Note that the local linear rate is also improved if $\theta_0$ is not the optimal choice, see \cite[chapter 5]{ciarlet1982introduction}. By remarking that for any choice of $\theta_0<2$ there is a neighborhood of $1$ on which $\varphi_{\theta_0}(\cdot)$ is nonnegative, one sees that this result applies in particular to our algorithm because eventually, for $\ell$ big enough, one has $\Theta(\gamma^\ell)=\theta_0$.
\begin{corollary}
The preceding local convergence rate applies to Algorithm \ref{thm:algo}, where $\theta_0$ is the chosen ``target'' overrelaxation parameter.
\end{corollary}
\begin{proof}
For this proof we will adopt a dual point of view, where the overrelaxed projection method boils down to the well-known SOR method~\eqref{SORdual}  and relying on the characterization $\gamma^{\ell}=e^{\alpha^\ell/\epsilon}\gamma^0e^{\beta^\ell/\epsilon}$ where $(\alpha^\ell,\beta^\ell)$ are the SOR iterates, initialized by $(\alpha^0,\beta^0)=(0,0)$. Notice that the dual problem~\eqref{DROT} has an invariant subspace due to the invariance by translation $(\alpha,\beta)\mapsto (\alpha-k,\beta+k)$, $k\in \IR$, but is strictly convex in any subspace which does not contain this line. Also, this invariance disappears in the associated primal iterates.

The SOR procedure~\eqref{SORdual}, when applied to maximize a concave quadratic form, defines a linear map $M_\omega : (\alpha^{\ell},\beta^{\ell})\mapsto (\alpha^{\ell+1},\beta^{\ell+1})$ which spectral properties are well known~\cite{ciarlet1982introduction, young2014iterative} , even in the non strictly concave case~\cite{hadjidimos1985optimization}, where convergence is replaced by \emph{semi-convergence}, i.e.\ convergence to a nonzero linear map.
